---
title: |
  | The Myth of the Myth of Independence:
  | A Critique of Binder and Spindel's Appraisal of the Fed
author: |
  | Zachary Thomas McDowell
  | BA & MA in Political Science
date: "**Updated**: `r format(Sys.time(), '%B %d, %Y')`"
abstract: |
  \singlespace The response to Desmond King's review of *The Myth of Independence* by one of the book's authors reveals one of two possible critical flaws in their thesis that the relationship between Congress and the Federal Reserve is best described as one of interdependence. Binder fails to adequately address the most important of King's critiques. This article expands on the points raised by King in a more comprehensive manner than a book review allows for, with a particular emphasis on contributing novel empirical analyses that cast further doubt on the arguments positied by Binder.
output:
  pdf_document:
    number_sections: true
    fig_caption: true
    keep_tex: TRUE
bibliography: monetary.bib
csl: apsa-edit.csl
fontsize: 11pt
indent: yes
link-citations: yes
geometry: margin=1in
header-includes:
- \usepackage{dcolumn}
- \usepackage{float}
- \floatplacement{figure}{ht}
- \usepackage{setspace}
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \usepackage{lastpage}
- \setlength{\headheight}{14.49998pt}
- \addtolength{\topmargin}{-2.49998pt}
- \usepackage{rotating, graphicx}
- \renewcommand{\topfraction}{.85}
- \renewcommand{\bottomfraction}{.7}
- \renewcommand{\textfraction}{.15}
- \renewcommand{\floatpagefraction}{.66}
- \setcounter{topnumber}{3}
- \setcounter{bottomnumber}{3}
- \setcounter{totalnumber}{4}
- \usepackage[hang,flushmargin]{footmisc}
- \usepackage{lipsum}
- \usepackage{etoolbox}
- \usepackage{mlmodern}
- \usepackage[T1]{fontenc}
- \usepackage{enumitem}
- \setlist{noitemsep}
- \AtBeginEnvironment{quote}{\singlespace\vspace{-\topsep}\small}
- \AtEndEnvironment{quote}{\vspace{-\topsep}\endsinglespace}
- |
    ```{=latex}
    \providecommand{\keywords}[1]{\textbf{Keywords:} #1}
    ```
nocite: |
  | @spindel2017, @king2018b, @lapavitsas2013, @lapavitsas2013c, @jacobs2016, @toussaint2019, @dorr2018a,
  @piketty2017, @piketty2020, @piketty2022, @boushey2018, @worldbank2022
---

```{r setup, include=FALSE}
#---------------------------------------------------------------------
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.width = 7,
                      fig.asp = 0.7,
                      fig.align = "center",
                      fig.pos = "ht")
#---------------------------------------------------------------------
library(tidyverse)
library(plm)
library(stargazer)
library(lmtest)
library(sandwich)
library(kableExtra)
library(interactions)
library(readxl)
library(captioner)
library(bookdown)
#---------------------------------------------------------------------
# Load Data
#---------------------------------------------------------------------
mega_combined_vars_df_final <- readRDS("~/Google Drive/My Drive/3 - Misc. Data Research/Edited Data/mega_combined_vars_df_final.rds")
#---------------------------------------------------------------------
fedr_df_v2_copy <- readRDS("~/Google Drive/My Drive/3 - Misc. Data Research/Edited Data/fedr_df_v2_copy.rds")
#---------------------------------------------------------------------
```

```{=latex}
\keywords{Inequality, Federal Reserve, Monetary Politics, Financial Crisis}
```

\let\thefootnote\relax\footnotetext{\lipsum[1][1-7]}

\thispagestyle{empty}

\newpage
\fancypagestyle{TOC}{%
    \fancyhead[R]{}
    \fancyhead[L]{}
}

\thispagestyle{TOC}
\tableofcontents
\listoffigures
\listoftables
\clearpage
\newpage

\fancypagestyle{INTRODUCTION}{%
    \fancyhead[R]{}
    \fancyhead[L]{}
}

\thispagestyle{INTRODUCTION}
\clearpage

\spacing{1.5}

# Introduction {-}

The politics of monetary policy is not the most popular subfield in political science. Therefore, it was curious to see that in 2018, the journal *Perspectives on Politics* published an interesting back-and-forth between two opposing sets of monetary politics scholars. Sarah Binder and Mark Spindel are the authors of *The Myth of Independence*, which argues that the Federal Reserve has an "interdependent relationship" with Congress. Both institutions rely on each other for different things, so neither is wholly dependent on nor independent from the other, hence why Binder and Spindel refer to their relationship as interdependent. On the other hand, Desmond King and Lawrence Jacobs argue in their book, *Fed Power*, that the Federal Reserve is effectively independent of Congress and holds more institutional power than it has at any point in its history. They argue that the Fed's policies, under normal economic conditions, serve to worsen economic inequality by favoring financial capital over working-class Americans. The *Perspectives on Politics* article is unique in that it is structured as follows:

1. King reviews *The Myth of Independence*
2. Binder and Spindel respond to King's review
3. Binder and Spindel reviews *Fed Power*
4. King and Lawrence respond to Binder and Spindel's review

In their response, King and Lawrence summarize their main disagreements with *The Myth of Independence*, stating that "Sarah Binder and Mark Spindel view the Fed’s actions as part of a dance with Congress that is largely silent about the winners and losers outside of Washington. By contrast, *Fed Power* puts the distributional consequences of the central bank’s policy front and center, along with the politics that produces them" [@binder2018c, pg. 783]. In other words, King criticizes Binder and Spindel for not paying enough attention to the distributional consequences concerning income and wealth that result from Fed policies, nor do they appear to recognize the actual amount of power held by the Fed. 

It is notable that these respective teams of Fed scholars would come to such diametrically opposed positions when it comes to the true nature of the relationship between Congress and the Fed. On its face, the question of whether or not the Fed is dependent on Congress would seem rather straightforward. In reality, as all four scholars point out, the quasi-governmental status of the Fed makes it difficult to study, given that its semi-private status prevents much information from being public accessible. But with that being said, there is enough public information available to where one might reasonably expect professional academics to be able to come to an agreement on a simple question like: What's the relationship between Congress and the Fed? 

# Initial Issues

Moreover, King and Lawrence, in their joint response, list their top four criticisms of *The Myth of Independence*,  wherein they posit that Binder and Spindel:

1. Overstate the Fed's "deference to Congress and underappreciates the Fed’s will and capacity to evade legislative control" [@binder2018c, pg. 783].
2. Underappreciate that the Fed "is independent of the congressional budget appropriations process" [@binder2018c, pg. 783].
3. Omit any discussion of the phenomenon known as "financialization" [@binder2018c, pg. 783].
4. Fail to recognize the clear institutional bias towards the wealthy, which can be seen in how "the Fed’s selective beneﬁts for ﬁnance and the conduct of monetary policy produce clear winners among the most afﬂuent" [@binder2018c, pg. 783].

What bothers me, as a scholar, is how Binder and Spindel egregiously failed to refute King and Lawrence on the actual terms he presented, almost as if they were intentionally ignoring the crux of King's criticisms because they had no way to counter him directly, instead relying on obfuscation to defend their claims. Binder and Spindel responded in such a way as to make it seem that they were avoiding the critiques raised in a seemingly condescending way.

Before we can begin arguing about whether or not the Fed is dependent on Congress, we must first define what we mean by "dependent." Further, we must also pin down the contextual and effective meaning of "accountability" when discussing the relationship between the two institutions. In my view, passing legislation is but one part of the matrix of components that constitutes what is necessary, but not sufficient, to hold the Fed accountable. I'm more interested in investigating whether or not it can be reasonably argued that the Fed, by virtue of its dependency on Congress, has been *meaningfully* held accountable by Congress for its various failings.

Without descending into a pedantic discussion of how to best understand and define accountability in its totality, my argument relies on the incorporation of one critical element involved in an entity being held accountable: for something to be held accountable, there must be at least some longevity in the newly applied constraints. In the context of the Fed, this would mean that any reforms passed by Congress would effectively constrain a particular set of dissatisfying actions taken by the Fed for the indefinite future, up to and until Congress legally undoes such reforms.

# Stasis Issues

The authors in question are not addressing each other, thus creating stasis issues in this "critical dialogue." It is frustrating to watch accomplished and no doubt knowledgeable Fed scholars talk past each other, but here we are. What follows is a systematic breakdown of the myriad stasis issues I identified in my reading of the critical back-and-forths presented in @binder2018c.

## Inequality & the Fed

The most egregious example of Binder and Spindel missing the point is how they respond to King's assertion that the Fed's post-2008 policy led to higher levels of income inequality. To highlight the degree to which they miss King's point, here are their respective words:

> **King**: Undeniably, the Fed’s interventions after the 2008 Great Recession to prevent a collapse of the ﬁnancial system in the United States and globally spared many from job loss and misery... it is important to avoid a false equivalency between the gains for ﬁnance and those for the general public. The Fed’s policies delivered lopsided and often concealed beneﬁts for ﬁnance, the top 1%, and the institutional interests of the Federal Reserve Bank that enjoys more power and autonomy than at any time in its 100-year history [@binder2018c, pg. 780].

> **Binder & Spindel**: Congress authorizes the Fed to make emergency loans to banks—not to steer aid directly to homeowners. In a crisis, monetary policy can affect the real economy by pumping credit through the clogged plumbing of the ﬁnancial system. And evidence from progressive economists suggests that the Fed’s unconventional bond purchases reduced mortgage rates, making working- and middle-class Americans better-off [@binder2018c, pg. 781].

I've done my best to fairly and accurately represent each author's views while being concise. If anyone has issues with how I've represented the above exchange, I welcome any criticisms, given the subjective nature of how I chose to shorten the respective quotes. That said, insofar as the above quotes *do* accurately and fairly represent the authors' respective views concerning the Fed's role in contributing to the 2008 financial crisis and the subsequent worsening of income and wealth inequality, on top of the aforementioned job losses and misery, Binder's response is lacking, to say the least.

Pointing to the fact that Fed's bond buying supposedly "reduced mortgage rates, making working- and middle-class Americans better-off" is tantamount to lauding an engineer for filling a single crack on a fully broken dam. Binder's response implicitly under-emphasizes the apocalyptic experiences felt by millions of Americans and citizens worldwide. Thomas Piketty is arguably the premier scholar when it comes to income and wealth inequality. In his book, *Capital in the Twenty-First Century*, Piketty argues two main, overarching points. First, the fundamental contradiction of capitalism is that, many times throughout the 18th to 21st centuries, the rate of return on capital assets was greater than the rate of economic growth, which he illustrates with the following simple expression: *r* > *g*.  @albertus2016a argues that, while Piketty's explanation is mostly accurate, the U-shaped pattern of inequality "only really holds for a handful of industrialized economies and a subset of developing countries" (pg. 49). Instead, the authors argue that the observed pattern of global income and wealth inequality seen in the 20th century resulted from the "political regime types and the social groups they empower, rather than war and globalization" [@albertus2016a, pg. 49].

There may not be enough time in this article to discuss the Fed's role in propping up the European Union during the Eurocrisis, but it's at least worth mentioning for the time being. For now, it is necessary to specify what we mean when discussing and subsequently operationalizing "economic inequality."

The most common way scholars measure *income* inequality is by calculating a country's Gini coefficient, an index value ranging from 0 to 1, with 0 representing perfect equality and 1 representing perfect inequality. The World Bank provides the following more comprehensive explanation:

> The Gini index measures the extent to which the distribution of income (or, in some cases, consumption expenditure) among individuals or households within an economy deviates from a perfectly equal distribution... The Gini index measures the area between the Lorenz curve and a hypothetical line of absolute equality, expressed as a percentage of the maximum area under the line. Thus a Gini index of 0 represents perfect equality, while an index of 100 implies perfect inequality [@worldbank2022].

\autoref{income_dist1} shows that, contrary to what Binder and Spindel seem to imply, inequality in the US worsened post-2008. And to bolster King and Jacobs's argument that the Fed should have emulated the Bank of Canada, \autoref{income_dist1} also shows that the income share of their Top 1% decreased sharply from 2008 to 2013. In contrast, the income share going to the Top 1% in the US  sharply *increased* post-2008. But where does the Fed come in? According to Leonard---and in stark contrast to the claims of Binder and Spindel---the Fed's policy of "quantitative easing" (QE) did more than any other single policy "to widen the divide between the rich and the poor [@leonard2022, pg. 10].

To critique King, he could have done a better job of demonstrating his claims concerning the Fed and the 2008 financial crisis by providing citations and at least a few more lines of discussion. But it is unclear if such improvements on King's part would have elicited a different response from Binder, given that the crux of their disagreement has nothing to do with the factual validity of each others' claims. Instead, they are discussing two completely different aspects of the financial crisis that can simultaneously be true. My issue with Binder's response has nothing to do with the factualness of her claim, rather that she 1) underplays the severity of the crisis and 2) makes it seem that the Fed's bond-buying *reduced* inequality rather than increased it.

It is also worth pointing out that the Fed's chosen vehicles for intervening in the economy (capital markets, bond markets, etc.) are not neutral in their distributional consequences. In other words, income inequality is partly fueled by the Fed's usage of various markets as the means to allocate the growing number of different financial assets that it owns and wants to inject into the broader economy. This is why it is said that any affect on the national economy caused by the Fed is indirect, since it has to first interact with the banking industry with the reasonable expectation that the banks' behavior will change in some vauge proportion to the benefit it received from the Fed.

Having recogized the distributional bias of Fed policy implementation, the next question we should ask ourselves is: Who benefits from the Fed operating this way? The answer: the very wealthy, those who already own assets. The Fed, in its role as a quasi-governmental institution, "oeprates in areas in which the wealthy possess unique, cumulative advantages" [@jacobs2018d, pg. 730]. 

## Independence vs. Interdependence

The thing that jumps out from Binder's response to King is their remarkable ignorance, whether it is willful or genuine, of the core content of King's critiques, especially on the topics of income inequality and the 2007-08 financial crisis. Let me be clear so scholars like Binder and Spindel have no excuse for misunderstanding or misrepresenting my position: The Federal Reserve is fully independent in all meaningful areas when it comes to its congressionally mandated responsibilities. I disagree with Binder and Spindel's argument. In my view, simply observing that there have been hardly more than a handful of instances since 1913 that Congress has "stepped in" to recalibrate the Fed is not sufficient to confidently assert that the Fed is, to a large degree, *dependent* on Congress due to a myriad of political concerns held by legislators. Rather, concurring with one of King's critiques, Binder and Spindel completely ignore the obvious fact that the Fed lobbies Congress for all sorts of things, not dissimilar to the behavior of private firms and special interest groups. By ignoring the mere capacity for active lobbying by the Fed, Binder and Spindel bias their observations in their favor. They treat every legislative action that concerns the Fed as external to and uninfluenced by the Fed itself, reinforcing Binder and Spindel's idea that the Fed is a mere subject of Congress and not an institution that frequently interacts and coordinates with Congress to achieve its own goals.

In my view, again concurring with King, *the* critical deficiency in Binder and Spindel's analysis is their frequent retreat "to a troubling form of argument about an 'intuitive' (p. 237) sense that the Fed is dependent on Congress, rather than a cogent demonstration of this proposition" [@king2018b, pg. 780]. No amount of pontificating about the very real pieces of legislation that have changed the parameters of the Fed's institutional capacity as a central bank is sufficient to demonstrate that the Fed truly, in any meaningful way, is *dependent* on Congress. Instead, they seem to find it sufficiently compelling to draw the reader's attention to the mere fact that Congress has passed laws that, on their face, succeed in holding the Fed *accountable*. But how, with a straight face, can someone say that the Fed has been effectively reigned in by legislation and successfully held accountable by Congress? As stated earlier, the Fed is the unequivocal most powerful central bank in the world, despite its history of drawing the ire of American voters, their elected representatives, and untold millions of others around the world that have felt, and will nevertheless continue feel, the effects of Fed policy. 

King cites Donald Kettl’s *Leadership at the Fed* as a complement to Binder and Spindel's book, wherein both works discuss the historical dysfunctionality of Congress as an institution. That said, King asks: How are we to believe that such a dysfunctional institution as Congress can regulate the Fed in any *meaningful* sense? Further, how can it be said that the Fed has been effectively held accountable when it is more powerful now than at any point in its history?

The main point is that Congress is barely capable of handling mundane legislative business, so it's a bit of a stretch to give it the benefit of the doubt concerning its capacity to control the Fed effectively. Could it be any other way? The Fed is more technically advanced in its operations and communications than Congress. Just look at the difference in budgets between the two institutions, and it will become clear which one holds the advantage when intellectually combating the other.

## Transparency vs. Accountability

Binder appears to equate legislation requiring increased transparency at the Fed as verifiable proof that Congress is fully capable of holding the Fed accountable for its actions, especially its failures. To pull from an earlier quote by King, the Fed "enjoys more power and autonomy than at any time in its 100-year history" [@binder2018c]. So the question we should be asking ourselves is: How can it be said that the Fed has been held accountable by Congress throughout its 100-year existence despite it currently wielding more power than ever? How does an institution accrue such power while being under the control of Congress?

## A People's History of the Fed

The scholarly treatment of the history of the Fed is one of the most curious phenomena in the social sciences. As Binder and Spindel did in *The Myth of Independence*, and as countless others have done and will continue to do, the Fed's essential history is well documented. That said, the *essential* history of the Fed is one of the obstacles that prevents most scholars from engaging with the Fed in recognition of what it is, that being the single most powerful and influential central bank in the world, an institution that garners the attention of both the finance ministers of every country as well as every bank and financial asset manager.

One need not look that far back into 20th-century history to observe the Fed's actions affecting the world economy, not to mention that one could argue that the European Union owes its post-2015 existence to the Fed's dollar swap-lines.

# Empirical Analysis

What causes inequality? More specifically, what causes income inequality? According to Binder and Spindel, not the Fed! Not the Fed's standard operating procedures! I'm more inclined to believe King and Jacobs when they say, on the contrary, that the Fed is an inequality machine, with each of its typical policies and actions serving to worsen inequality in the US and the world. But which set of scholars is right?

To see which book's thesis holds up better under scrutiny, I have created an empirical model designed to capture all of the major contributors to income inequality. The first question of interest concerns the role of financial capital assets in worsening income inequality.

## Literature Review

There is a large body of literature that examines the various social and political phenomena caused by inequality, but less work has been done to explain what causes inequality in the first place.

## Hypotheses

-   **H1**: As financial assets *increase* as a share of GDP, income inequality will *increase.*

-   **H2**: As union density *increases*, income inequality will *decrease.*

-   **H3**: As the real rate of return on capital *increases*, inequality will *increase.*

## Methods

Following the recommendations of Nathaniel Beck and Jonathan Katz, I have employed a time-series panel study of 24 OECD countries that applies panel-corrected standard errors (PCSE) in the final analysis.

## Results

\autoref{ineq1} shows, across multiple model specifications, that a one unit increase in financial assets (as a share of GDP) is associated with, on average and all things being equal, between a 0.204 and 1.346 unit increase in income inequality, depending on how inequality is measured. The main takeaway from this table is that a number of things appear to have an effect on income inequality, regardless of how it is measured.

# Conclusion

Binder and Spindel fail to effectively refute any of King and Jacobs's criticisms. Instead, they reinforce the validity of the various criticisms by intentionally refusing or failing to provide any comprehensive counter-arguments, only ones that touch on one particular *aspect* of the original criticism.  Thus, it is difficult, if not impossible, to find Binder and Spindel's thesis credible when they have shown that they are either unwilling or incapable of defending their claims. 

\clearpage
\newpage
\onehalfspace
\fancypagestyle{BIB}{%
    \fancyhead[R]{}
    \fancyhead[L]{}
}

\thispagestyle{BIB}

# Bibliography {-}

<div id="refs"></div>

\singlespace
\newpage
\fancypagestyle{APP}{%
    \fancyhead[R]{}
    \fancyhead[L]{}
}

\pagestyle{APP}

# Appendix {-}


```{r ols1, cache=TRUE, results='asis'}
#--------------------------------------------------------------------------------------
regression_ready_df <- mega_combined_vars_df_final %>%
  filter(year >= 1980) %>%
  select(iso3, year,
         wid_pre_tax_income_share_top_1pct,
         wid_ratio_income_share_99th_50th_pctile,
         wid_ratio_income_share_99th_10th_pctile,
         wid_ratio_income_share_90th_50th_pctile,
         wid_ratio_income_share_90th_10th_pctile,
         pwt_irr,
         cpds_unemployment_rate,
         cpds_inflation_rate,
         oecd_union_density,
         pwt_gdp_pc_growth_2017_USD,
         pwt_human_capital_index,
         wid_pct_gdp_national_financial_assets,
         pwt_pct_gdp_international_trade,
         cabinet_majority_left,
         oecd_public_socspend_wa_pct_gdp) %>%
  make.pbalanced()
#-----------------------------------------------------------------------------------------------
form_m1.1 <- wid_pre_tax_income_share_top_1pct ~
  pwt_irr +
  cpds_unemployment_rate +
  cpds_inflation_rate +
  oecd_union_density +
  pwt_gdp_pc_growth_2017_USD +
  factor(iso3)

form_m2.1 <- wid_pre_tax_income_share_top_1pct ~
  pwt_irr +
  cpds_unemployment_rate +
  cpds_inflation_rate +
  oecd_union_density +
  pwt_gdp_pc_growth_2017_USD +
  pwt_human_capital_index +
  wid_pct_gdp_national_financial_assets +
  pwt_pct_gdp_international_trade +
  oecd_public_socspend_wa_pct_gdp +
  cabinet_majority_left +
  factor(iso3)

form_m1.2 <- wid_ratio_income_share_99th_50th_pctile ~
  pwt_irr +
  cpds_unemployment_rate +
  cpds_inflation_rate +
  oecd_union_density +
  pwt_gdp_pc_growth_2017_USD +
  factor(iso3)

form_m2.2 <- wid_ratio_income_share_99th_50th_pctile ~
  pwt_irr +
  cpds_unemployment_rate +
  cpds_inflation_rate +
  oecd_union_density +
  pwt_gdp_pc_growth_2017_USD +
  pwt_human_capital_index +
  wid_pct_gdp_national_financial_assets +
  pwt_pct_gdp_international_trade +
  oecd_public_socspend_wa_pct_gdp +
  cabinet_majority_left +
  factor(iso3)

form_m1.3 <- wid_ratio_income_share_99th_10th_pctile ~
  pwt_irr +
  cpds_unemployment_rate +
  cpds_inflation_rate +
  oecd_union_density +
  pwt_gdp_pc_growth_2017_USD +
  factor(iso3)

form_m2.3 <- wid_ratio_income_share_99th_10th_pctile ~
  pwt_irr +
  cpds_unemployment_rate +
  cpds_inflation_rate +
  oecd_union_density +
  pwt_gdp_pc_growth_2017_USD +
  pwt_human_capital_index +
  wid_pct_gdp_national_financial_assets +
  pwt_pct_gdp_international_trade +
  oecd_public_socspend_wa_pct_gdp +
  cabinet_majority_left +
  factor(iso3)
#-----------------------------------------------------------------------------------------------
# Models
#-----------------------------------------------------------------------------------------------
lm_m1 <- lm(form_m1.1, data = regression_ready_df)
lm_m2 <- lm(form_m2.1, data = regression_ready_df)
lm_m3 <- lm(form_m1.2, data = regression_ready_df)
lm_m4 <- lm(form_m2.2, data = regression_ready_df)
lm_m5 <- lm(form_m1.3, data = regression_ready_df)
lm_m6 <- lm(form_m2.3, data = regression_ready_df)
#-----------------------------------------------------------------------------------------------
# Robust Standard Error Objects
#-----------------------------------------------------------------------------------------------
hc_se_lm_m1 <- sqrt(diag(vcovHC(lm_m1, type = "HC1")))
hc_se_lm_m2 <- sqrt(diag(vcovHC(lm_m2, type = "HC1")))
hc_se_lm_m3 <- sqrt(diag(vcovHC(lm_m3, type = "HC1")))
hc_se_lm_m4 <- sqrt(diag(vcovHC(lm_m4, type = "HC1")))
hc_se_lm_m5 <- sqrt(diag(vcovHC(lm_m5, type = "HC1")))
hc_se_lm_m6 <- sqrt(diag(vcovHC(lm_m6, type = "HC1")))
#-----------------------------------------------------------------------------------------------
# PCSE Standard Error Objects
#-----------------------------------------------------------------------------------------------
pcse_se_lm_m1 <- sqrt(diag(vcovPC(lm_m1, cluster = ~ iso3 + year)))
pcse_se_lm_m2 <- sqrt(diag(vcovPC(lm_m2, cluster = ~ iso3 + year)))
pcse_se_lm_m3 <- sqrt(diag(vcovPC(lm_m3, cluster = ~ iso3 + year)))
pcse_se_lm_m4 <- sqrt(diag(vcovPC(lm_m4, cluster = ~ iso3 + year)))
pcse_se_lm_m5 <- sqrt(diag(vcovPC(lm_m5, cluster = ~ iso3 + year)))
pcse_se_lm_m6 <- sqrt(diag(vcovPC(lm_m6, cluster = ~ iso3 + year)))
#-----------------------------------------------------------------------------------------------
# Creating List of LM Objects
#-----------------------------------------------------------------------------------------------
lm_group <- list(lm_m1, lm_m2, lm_m3, lm_m4, lm_m5, lm_m6)
#-----------------------------------------------------------------------------------------------
hc.se_group <- list(hc_se_lm_m1, hc_se_lm_m2, hc_se_lm_m3, hc_se_lm_m4, hc_se_lm_m5, hc_se_lm_m6)
#-----------------------------------------------------------------------------------------------
pcse.se_group <- list(pcse_se_lm_m1, pcse_se_lm_m2, pcse_se_lm_m3, pcse_se_lm_m4, pcse_se_lm_m5, pcse_se_lm_m6)
#-----------------------------------------------------------------------------------------------
stargazer(lm_group,
          type = "latex",
          se = c(pcse.se_group),
          omit = c("factor\\(iso3\\)"),
          keep.stat = c("n", "rsq", "adj.rsq"),
          title = "\\label{ineq1}Determinants of Income Inequality (FE)",
          dep.var.labels.include = FALSE,
          model.names = FALSE,
          header = FALSE,
          no.space = TRUE,
          float = TRUE,
          float.env = "table",
          font.size = "small",
          covariate.labels = c("IRR",
                               "Unemployment",
                               "Inflation",
                               "Union Density",
                               "GDP Growth",
                               "Human Capital",
                               "Financial Assets",
                               "Trade",
                               "Welfare",
                               "Left Cabinet"))
```

```{r ols2, cache=TRUE, results='asis'}
#--------------------------------------------------------------------------------------
regression_ready_df <- mega_combined_vars_df_final %>%
  filter(year >= 1980) %>%
  select(iso3, year,
         wid_pre_tax_income_share_top_1pct,
         wid_ratio_income_share_99th_50th_pctile,
         wid_ratio_income_share_99th_10th_pctile,
         wid_ratio_income_share_90th_50th_pctile,
         wid_ratio_income_share_90th_10th_pctile,
         pwt_irr,
         cpds_unemployment_rate,
         cpds_inflation_rate,
         oecd_union_density,
         pwt_gdp_pc_growth_2017_USD,
         pwt_human_capital_index,
         wid_pct_gdp_national_financial_assets,
         pwt_pct_gdp_international_trade,
         cabinet_majority_left,
         oecd_public_socspend_wa_pct_gdp) %>%
  make.pbalanced()
#-----------------------------------------------------------------------------------------------
form_m1.1 <- wid_pre_tax_income_share_top_1pct ~
  pwt_irr +
  cpds_unemployment_rate +
  cpds_inflation_rate +
  oecd_union_density +
  pwt_gdp_pc_growth_2017_USD

form_m2.1 <- wid_pre_tax_income_share_top_1pct ~
  pwt_irr +
  cpds_unemployment_rate +
  cpds_inflation_rate +
  oecd_union_density +
  pwt_gdp_pc_growth_2017_USD +
  pwt_human_capital_index +
  wid_pct_gdp_national_financial_assets +
  pwt_pct_gdp_international_trade +
  oecd_public_socspend_wa_pct_gdp +
  cabinet_majority_left

form_m1.2 <- wid_ratio_income_share_99th_50th_pctile ~
  pwt_irr +
  cpds_unemployment_rate +
  cpds_inflation_rate +
  oecd_union_density +
  pwt_gdp_pc_growth_2017_USD

form_m2.2 <- wid_ratio_income_share_99th_50th_pctile ~
  pwt_irr +
  cpds_unemployment_rate +
  cpds_inflation_rate +
  oecd_union_density +
  pwt_gdp_pc_growth_2017_USD +
  pwt_human_capital_index +
  wid_pct_gdp_national_financial_assets +
  pwt_pct_gdp_international_trade +
  oecd_public_socspend_wa_pct_gdp +
  cabinet_majority_left

form_m1.3 <- wid_ratio_income_share_99th_10th_pctile ~
  pwt_irr +
  cpds_unemployment_rate +
  cpds_inflation_rate +
  oecd_union_density +
  pwt_gdp_pc_growth_2017_USD

form_m2.3 <- wid_ratio_income_share_99th_10th_pctile ~
  pwt_irr +
  cpds_unemployment_rate +
  cpds_inflation_rate +
  oecd_union_density +
  pwt_gdp_pc_growth_2017_USD +
  pwt_human_capital_index +
  wid_pct_gdp_national_financial_assets +
  pwt_pct_gdp_international_trade +
  oecd_public_socspend_wa_pct_gdp +
  cabinet_majority_left
#-----------------------------------------------------------------------------------------------
# Models
#-----------------------------------------------------------------------------------------------
lm_m1 <- lm(form_m1.1, data = regression_ready_df)
lm_m2 <- lm(form_m2.1, data = regression_ready_df)
lm_m3 <- lm(form_m1.2, data = regression_ready_df)
lm_m4 <- lm(form_m2.2, data = regression_ready_df)
lm_m5 <- lm(form_m1.3, data = regression_ready_df)
lm_m6 <- lm(form_m2.3, data = regression_ready_df)
#-----------------------------------------------------------------------------------------------
# Robust Standard Error Objects
#-----------------------------------------------------------------------------------------------
hc_se_lm_m1 <- sqrt(diag(vcovHC(lm_m1, type = "HC1")))
hc_se_lm_m2 <- sqrt(diag(vcovHC(lm_m2, type = "HC1")))
hc_se_lm_m3 <- sqrt(diag(vcovHC(lm_m3, type = "HC1")))
hc_se_lm_m4 <- sqrt(diag(vcovHC(lm_m4, type = "HC1")))
hc_se_lm_m5 <- sqrt(diag(vcovHC(lm_m5, type = "HC1")))
hc_se_lm_m6 <- sqrt(diag(vcovHC(lm_m6, type = "HC1")))
#-----------------------------------------------------------------------------------------------
# PCSE Standard Error Objects
#-----------------------------------------------------------------------------------------------
pcse_se_lm_m1 <- sqrt(diag(vcovPC(lm_m1, cluster = ~ iso3 + year)))
pcse_se_lm_m2 <- sqrt(diag(vcovPC(lm_m2, cluster = ~ iso3 + year)))
pcse_se_lm_m3 <- sqrt(diag(vcovPC(lm_m3, cluster = ~ iso3 + year)))
pcse_se_lm_m4 <- sqrt(diag(vcovPC(lm_m4, cluster = ~ iso3 + year)))
pcse_se_lm_m5 <- sqrt(diag(vcovPC(lm_m5, cluster = ~ iso3 + year)))
pcse_se_lm_m6 <- sqrt(diag(vcovPC(lm_m6, cluster = ~ iso3 + year)))
#-----------------------------------------------------------------------------------------------
# Creating List of LM Objects
#-----------------------------------------------------------------------------------------------
lm_group <- list(lm_m1, lm_m2, lm_m3, lm_m4, lm_m5, lm_m6)
#-----------------------------------------------------------------------------------------------
hc.se_group <- list(hc_se_lm_m1, hc_se_lm_m2, hc_se_lm_m3, hc_se_lm_m4, hc_se_lm_m5, hc_se_lm_m6)
#-----------------------------------------------------------------------------------------------
pcse.se_group <- list(pcse_se_lm_m1, pcse_se_lm_m2, pcse_se_lm_m3, pcse_se_lm_m4, pcse_se_lm_m5, pcse_se_lm_m6)
#-----------------------------------------------------------------------------------------------
stargazer(lm_group,
          type = "latex",
          se = c(pcse.se_group),
          omit = c("factor\\(iso3\\)"),
          keep.stat = c("n", "rsq", "adj.rsq"),
          title = "\\label{ineq2}Determinants of Income Inequality (Pooled)",
          dep.var.labels.include = FALSE,
          model.names = FALSE,
          header = FALSE,
          no.space = TRUE,
          float = TRUE,
          float.env = "table",
          font.size = "small",
          covariate.labels = c("IRR",
                               "Unemployment",
                               "Inflation",
                               "Union Density",
                               "GDP Growth",
                               "Human Capital",
                               "Financial Assets",
                               "Trade",
                               "Welfare",
                               "Left Cabinet"))
```

```{r fed_funds, cache=TRUE, fig.cap="\\label{fed_funds_rate}Effective Federal Funds Rate (\\%)"}
#--------------------------------------------------------------------------------
fedr_df_v2_copy %>%
  filter(year >= 1970) %>%
  filter(series_id %in% c('Federal Funds Rate')) %>%
  ggplot(mapping = aes(x = date, y = value)) +
  geom_line() +
  geom_smooth(method = "loess") +
  theme_bw() +
  labs(x = "Month-Year",
       y = "Interest Rate (%)") +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        axis.title.x = element_text(vjust = -1),
        text = element_text(face = 'bold'),
        legend.position = "bottom",
        legend.background = element_blank(),
        legend.box.background = element_rect(colour = "black")) +
  scale_x_date(date_labels = "%b-%Y",
               date_breaks = "48 month") +
  geom_vline(xintercept=as.numeric(fedr_df_v2_copy$date[299]), linetype="longdash", color = "red") +
  geom_label(label="Volcker Shock",
             x = as.numeric(fedr_df_v2_copy$date[237]),
             y = 18.5,
             label.padding = unit(0.275, "lines"),
             label.size = 0.15,
             color = "black",
             fill = "#69b3a2")
#--------------------------------------------------------------------------------
```

```{r income_dist1, cache=TRUE, fig.cap="\\label{income_dist1}Share of National Income (\\%)"}
#--------------------------------------------------------------------------------
mega_combined_vars_df_final %>%
  filter(iso2 %in% c("US", "CA")) %>%
  mutate(wid_pre_tax_income_share_bottom_50pct = wid_pre_tax_income_share_bottom_50pct*100,
         wid_pre_tax_income_share_top_1pct = wid_pre_tax_income_share_top_1pct*100) %>%
  rename("Bottom 50%" = wid_pre_tax_income_share_bottom_50pct,
         "Top 1%" = wid_pre_tax_income_share_top_1pct,) %>%
  select(country_name, year, "Bottom 50%", "Top 1%") %>%
  gather(key = "Type", value = "value", -country_name, -year) %>%
  ggplot(aes(x = year, y = value)) +
  geom_line(aes(linetype = Type)) +
  theme_bw() +
  labs(x = "Year",
       y = "Income Share (%)",
       linetype = "",
       caption = "
       Data source(s): World Inequality Database") +
  theme(text = element_text(face = 'bold'),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5),
        plot.caption = element_text(hjust = 0),
        plot.caption.position = "plot",
        legend.position = "bottom",
        legend.background = element_blank(),
        legend.box.background = element_rect(colour = "black")) +
  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 10)) +
  scale_y_continuous(breaks = seq(7, 22, by = 2)) +
  facet_wrap(vars(country_name)) +
  geom_vline(xintercept = 2008, linetype = "dashed", color = "blue")
#--------------------------------------------------------------------------------
```

```{r capital_comp, cache=TRUE, fig.cap="\\label{income_dist1}Cross-National Comparison of Capital Investments (\\% of GDP)"}
#--------------------------------------------------------------------------------
mega_combined_vars_df_final %>%
  filter(iso2 %in% c("US", "CA")) %>%
  select(country_name, year,
         wid_pct_gdp_national_financial_assets,
         wid_pct_gdp_national_nonfinancial_assets) %>%
  rename("Financial Assets" = wid_pct_gdp_national_financial_assets,
         "Nonfinancial Assets" = wid_pct_gdp_national_nonfinancial_assets) %>%
  gather(key = "Type", value = "value", -country_name, -year) %>%
  ggplot(aes(x = year, y = value)) +
  geom_line(aes(linetype = Type)) +
  theme_bw() +
  labs(x = "Year",
       y = "Percent GDP (%)",
       linetype = "",
       caption = "
       Data source(s): World Inequality Database & Penn World Table 10.0") +
  theme(text = element_text(face = 'bold'),
        axis.title.x = element_text(vjust = -0.75),
        plot.caption = element_text(hjust = 0),
        plot.caption.position = "plot",
        legend.position = "bottom",
        legend.background = element_blank(),
        legend.box.background = element_rect(colour = "black")) +
  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 10)) +
  geom_vline(xintercept = 2007, linetype = "dashed", color = "blue") +
  facet_wrap(vars(country_name))
```
