---
title: |
  | The Myth of the Myth of Independence:
  | A Critique of Binder and Spindel's Appraisal of the Fed
author: |
  | Zachary Thomas McDowell
  | BA & MA in Political Science 
  | University of Georgia
date: "**Updated**: `r format(Sys.time(), '%B %d, %Y')`"
abstract: |
  \singlespace The response to Desmond King's review of *The Myth of Independence* by one of the book's authors reveals one of two possible critical flaws in their thesis that the relationship between Congress and the Federal Reserve is best described as one of interdependence. Binder fails to adequately address the most important of King's critiques. This article expands on the points raised by King in a more comprehensive manner than a book review allows for, with a particular emphasis on contributing novel empirical analyses that cast further doubt on the arguments positied by Binder.
output:
  pdf_document:
    number_sections: true
    fig_caption: true
bibliography: monetary.bib
csl: apsa-edit.csl
fontsize: 12pt
indent: yes
link-citations: yes
geometry: margin=1in
header-includes:
- \usepackage{dcolumn}
- \usepackage{float}
- \floatplacement{figure}{ht}
- \usepackage{setspace}
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \usepackage{lastpage}
- \setlength{\headheight}{14.49998pt}
- \addtolength{\topmargin}{-2.49998pt}
- \usepackage{rotating, graphicx}
- \renewcommand{\topfraction}{.85}
- \renewcommand{\bottomfraction}{.7}
- \renewcommand{\textfraction}{.15}
- \renewcommand{\floatpagefraction}{.66}
- \setcounter{topnumber}{3}
- \setcounter{bottomnumber}{3}
- \setcounter{totalnumber}{4}
- \usepackage[hang,flushmargin]{footmisc}
- \usepackage{lipsum}
- \usepackage{etoolbox}
- \usepackage{mlmodern}
- \usepackage[T1]{fontenc}
- \AtBeginEnvironment{quote}{\singlespace\vspace{-\topsep}\small}
- \AtEndEnvironment{quote}{\vspace{-\topsep}\endsinglespace}
- |
    ```{=latex}
    \providecommand{\keywords}[1]{\textbf{Keywords:} #1}
    ```
nocite: |
  | @spindel2017, @king2018b, @lapavitsas2013, @lapavitsas2013c, @jacobs2016, @dorr2018a, @szepanski2022, @toussaint2019, @dorr2018a,
  @piketty2017, @piketty2020, @piketty2022, @boushey2018, @axilrod2011, @tooze2018, @crisisrep2011
---

```{r setup, include=FALSE}
#---------------------------------------------------------------------
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.width = 6.25,
                      fig.asp = 0.8,
                      fig.align = "center",
                      fig.pos = "ht")
#---------------------------------------------------------------------
library(tidyverse)
library(stargazer)
library(lmtest)
library(sandwich)
library(kableExtra)
library(interactions)
library(readxl)
library(captioner)
library(bookdown)
#---------------------------------------------------------------------
# Load Data
#---------------------------------------------------------------------
mega_combined_vars_df <- readRDS("~/Google Drive/My Drive/3 - Misc. Data Research/Fed Data/mega_combined_vars_df.rds")
#---------------------------------------------------------------------
new_mega_df <- readRDS("~/Google Drive/My Drive/3 - Misc. Data Research/Fed Data/new_mega_df.rds")
#---------------------------------------------------------------------
new_mega_edits_df <- readRDS("~/Google Drive/My Drive/3 - Misc. Data Research/Fed Data/new_mega_edits_df.rds")
#---------------------------------------------------------------------
pwt_mega_df <- readRDS("~/Google Drive/My Drive/3 - Misc. Data Research/Fed Data/pwt_mega_df.rds")
#---------------------------------------------------------------------
fedr_df_v2_copy <- readRDS("~/Google Drive/My Drive/3 - Misc. Data Research/Fed Data/fedr_df_v2_copy.rds")
#---------------------------------------------------------------------
```

```{=latex}
\keywords{Inequality, Fed, Monetary Politics, Financial Crisis}
```

\let\thefootnote\relax\footnotetext{\lipsum[1][1-7]}

\thispagestyle{empty}

\newpage
\fancypagestyle{TOC}{%
    \fancyhead[R]{}
    \fancyhead[L]{}
}

\thispagestyle{TOC}
\tableofcontents
\listoffigures
\listoftables
\clearpage
\newpage

\fancypagestyle{INTRODUCTION}{%
    \fancyhead[R]{}
    \fancyhead[L]{}
}

\thispagestyle{INTRODUCTION}
\clearpage

\spacing{1.5}

# Introduction {-}

Man, that review response really bothered me. Like, the book's authors didn't even seem to try to refute King on the actual terms he presented, almost as if they there will intentionally ignoring the crux of King's criticisms because they had no way to counter him directly, instead relying on obfuscation to defend their claims. Sincerely, Binder and Spindel responded in such a way as make it seem that they were avoiding the critiques raised in an almost condescending way.

Before we can begin arguing about whether or not the Fed is dependent on Congress, we must first define what me mean by "dependent." Further, we must also pin down the contextual and effective meaning of "accountability" when discussing the relationship between the two institutions. In my view, passing legislation is but one part of the matrix of components that constitutes what is necessary, but not sufficient, to hold the Fed accountable. I'm more interested in investigating whether or not it can be reasonably argued that the Fed, by virtue of its dependency on Congress, has been *meaningfully* held accountable by Congress for its various failings.

Without descending into a pedantic discussion of how to best understand and define accountability in its totality, my argument relies on the incorporation of one critical element involved in an entity being held accountable: for something to be held accountable, there must be at least some longevity in the newly applied constraints. In the context of the Fed, this would mean that any reforms passed by Congress would effectively constrain a particular set of dissatisfying actions taken by the Fed for the indefinite future, up to and until such reforms are legally undone by Congress.

# Stasis Issues

It's obvious that the authors in question are not addressing each other, thus creating stasis issues in this "critical dialogue." It is frustrating to watch accomplished and no doubt knowledge Fed scholars talk past each other, but nonetheless here we are. What follows is a systematic breakdown of the myriad stasis issues I identified in my reading of the critical back-and-forths presented in @binder2018c.

## Inequality & the Fed

The most egregious example of Binder totally missing the point is how she responds to King's assertion that the Fed's post-2008 policy led to higher levels of income inequality. To highlight the degree to which Binder misses King's point, here are their respective words:

> **King**: Undeniably, the Fed’s interventions after the 2008 Great Recession to prevent a collapse of the ﬁnancial system in the United States and globally spared many from job loss and misery... it is important to avoid a false equivalency between the gains for ﬁnance and those for the general public. The Fed’s policies delivered lopsided and often concealed beneﬁts for ﬁnance, the top 1%, and the institutional interests of the Federal Reserve Bank that enjoys more power and autonomy than at any time in its 100-year history [@binder2018c, pg. 780].

> **Binder**: Congress authorizes the Fed to make emergency loans to banks—not to steer aid directly to homeowners. In a crisis, monetary policy can affect the real economy by pumping credit through the clogged plumbing of the ﬁnancial system. And evidence from progressive economists suggests that the Fed’s unconventional bond purchases reduced mortgage rates, making working- and middleclass Americans better-off [@binder2018c, pg. 781].

I've done my best to fairly and accurately represent each authors' views, while still being concise. If anyone has issues with how I've represented the above exchange, I welcome any and all criticisms, given the subjective nature of how I chose to shorten the respective quotes. That said, insofar as the above quotes *do* accurately and fairly represent the authors' respective views concerning the Fed's role in contributing to the 2008 financial crisis and the subsequent worsening of income and wealth inequality, on top of the aforementioned job losses and misery, Binder's response is pathetic, to say the least. Pointing to the fact that Fed's bond buying supposedly "reduced mortgage rates, making working- and middleclass Americans better-off" is tantamount to lauding an engineer for filling a single crack on a fully broken dam. Binder's response implicitly under-emphasizes the apocalyptic experiences felt by millions of Americans and citizens in countries the world over. There may not be enough time in this article to discuss the Fed's role in propping up the European Union during the Eurocrisis, but its at least worth mentioning for the time being.

```{r income_dist1, cache=TRUE, fig.cap="\\label{income_dist1}Share of National Income (\\%)"}
#--------------------------------------------------------------------------------
# Solid Barplot
#--------------------------------------------------------------------------------
mega_combined_vars_df %>%
  filter(country_code_ISO3 %in% c("USA")) %>%
  rename("Bottom 50%" = WID_pre_tax_income_share_bottom_50pct,
         "Top 1%" = WID_pre_tax_income_share_top_1pct) %>%
  select(country_name, year, "Bottom 50%", "Top 1%") %>%
  gather(key = "Type", value = "value", -country_name, -year) %>%
  ggplot(aes(x = year, y = value)) +
  geom_line(aes(linetype = Type)) +
  theme_bw() +
  labs(x = "Year",
       y = "Share of National Income (%)",
       linetype = "",
       caption = "
       Data source(s): World Inequality Database") +
  theme(text = element_text(face = 'bold'),
        plot.caption = element_text(hjust = 0, face = "bold"),
        plot.caption.position = "plot",
        legend.position = "bottom",
        legend.background = element_blank(),
        legend.box.background = element_rect(colour = "black")) +
  scale_x_continuous(limits = c(1990, 2020), breaks = seq(1990, 2020, by = 10))
#--------------------------------------------------------------------------------
```

To critique King, he could have done a better job of demonstrating his claims concerning the Fed and the 2008 financial crisis by providing citations and at least a few more lines of discussion. But it is unclear if such improvements on King's part would have elicited a different response from Binder, given that the crux of their disagreement has nothing to do with the factual validity of each others' claims. Instead, they are effectively discussing two completely different aspects of the financial crisis that can simultaneously be true. My issue with Binder's response has nothing to do with the factualness of her claim, rather that she 1) underplays the severity of the crisis and 2) makes it seem that the Fed's bond buying *reduced* inequality rather than increase it.

## Independence vs. Interdependence

The thing that jumps out from Binder's response to King is their remarkable ignorance, whether it is willful or genuine, of the core content of King's critiques, especially on the topics of income inequality and the 2007-08 financial crisis. Let me be clear, so that scholars like Binder and Spindel have no excuse for misunderstanding or misrepresenting my position: The Federal Reserve is fully independent in all meaningful areas when it comes to its congressionally mandated responsibilities. I disagree with Binder and Spindel's argument that simply observing that there are barely more than a handful of instances since 1913 when Congress has "stepped in" to recalibrate the Fed is sufficient evidence to confidently assert that the Fed is, to a large degree, *dependent* on Congress due to a myriad of political concerns held by legislators. Rather, concurring with one of King's critiques, Binder and Spindel completely ignore the very obvious fact that the Fed lobbies Congress for all sorts of things, not dissimilar to the behavior of private firms and special interest groups. By ignoring the mere capacity for active lobbying by the Fed, Binder and Spindel bias their observations in their favor. They treat every legislative action that concerns the Fed as totally external to and uninfluenced by the Fed itself, which serves to reinforce Binder and Spindel's idea that the Fed is a mere subject of Congress and not an institution that frequently interacts and coordinates with Congress to achieve its own goals.

In my view, again concurring with King, *the* critical deficiency in Binder and Spindel's analysis is their frequent retreat "to a troubling form of argument about an 'intuitive' (p. 237) sense that the Fed is dependent on Congress, rather than a cogent demonstration of this proposition" [@king2018b, pg. 780]. No amount of pontificating about the very real pieces of legislation that have changed the parameters of the Fed's institutional capacity as a central bank is sufficient to demonstrate that the Fed truly, in any meaningful way, is *dependent* on Congress. Instead, they seem to find it sufficiently compelling to draw the reader's attention to the mere fact that Congress has passed laws that, on their face, succeed in holding the Fed *accountable*. But how, with a straight face, can someone say that the Fed has been effectively reigned in by legislation, has been successfully held accountable by Congress? As stated earlier, the Fed is the unequivocal most powerful central bank in the world, despite its history of drawing the ire of American voters, their elected representatives, and untold millions of others around the world that have felt, and will nevertheless continue feel, the effects of Fed policy. 

King cites Donald Kettl’s *Leadership at the Fed* as a complement to Binder and Spindel's book, wherein both works discuss the historical dysfunctionality of Congress as an institution. That said, King asks: How are we to believe that such a dysfunctional institution as Congress is at all capable of regulating the Fed in any *meaningful* sense? Further, how can it be said that the Fed has been effectively held accountable when it is more powerful now than at any point in its history?

The main point to make is that Congress is barely capable of handling mundane legislative business, so its a bit of a stretch to give it the benefit of the doubt concerning its capacity to effectively control the Fed. Could it be any other way? The Fed is vastly more technically advanced in its operations and communications than Congress. Just look at the difference in budgets between the two institutions and it will become clear which one holds the advantage when it comes intellectually combating the other.

## Transparency vs. Accountability

Binder appears to equate legislation requiring increased transparency at the Fed as demonstrable proof that Congress is fully capable of holding the Fed accountable for its actions, especially its failures. To pull from an earlier quote by King, the Fed "enjoys more power and autonomy than at any time in its 100-year history" [@binder2018c]. So the question we should be asking ourselves is: How can it be said that the Fed has been held accountable by Congress throughout its 100 year existence despite it currently wielding more power than ever? How does an institution accrue such power while being under the control of Congress?

# A People's History of the Fed

The scholarly treatment of the history of the Fed is one of the most curious phenomena in the social sciences. As Binder and Spindel did in *The Myth of Independence*, and as countless others have done and will continue to do, the Fed's essential history is well documented. That said, the *essential* history of the Fed is one of the obstacles that prevents most scholars from engaging with the Fed in recognition of what it is, that being the single most powerful and influential central banks in the world, an institution that garners the attention of both the finance minsters of every country as well as every bank and financial asset manager.

```{r fed_funds, cache=TRUE, fig.cap="\\label{fed_funds_rate}Effective Federal Funds Rate (\\%)"}
#--------------------------------------------------------------------------------
fedr_df_v2_copy %>%
  filter(year >= 1970) %>%
  filter(series_id %in% c('Federal Funds Rate')) %>%
  ggplot(mapping = aes(x = date, y = value)) +
  geom_line() +
  geom_smooth(method = "loess") +
  theme_bw() +
  labs(x = "Month-Year",
       y = "Interest Rate (%)") +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        axis.title.x = element_text(vjust = -1),
        text = element_text(face = 'bold'),
        legend.position = "bottom",
        legend.background = element_blank(),
        legend.box.background = element_rect(colour = "black")) +
  scale_x_date(date_labels = "%b-%Y",
               date_breaks = "48 month") +
  geom_vline(xintercept=as.numeric(fedr_df_v2_copy$date[299]), linetype="longdash", color = "red") +
  geom_label(label="Volcker Shock",
             x = as.numeric(fedr_df_v2_copy$date[237]),
             y = 18.5,
             label.padding = unit(0.275, "lines"),
             label.size = 0.15,
             color = "black",
             fill = "#69b3a2")
#--------------------------------------------------------------------------------
```

One need not look far back into 20th century history to observe the Fed's actions affecting the world economy, not to mention the fact that one could argue that the European Union owes its post-2015 existence to the Fed's dollar swap-lines.  

# Conclusion

Binder fails to refute any of King's criticisms. Instead, she reinforces there validity, either by intentionally refusing or simply failing, to provide any comprehensive counter-arguments, only ones that touch on one particular aspect of the original criticism.

\clearpage
\newpage
\onehalfspace
\fancypagestyle{BIB}{%
    \fancyhead[R]{}
    \fancyhead[L]{}
}

\thispagestyle{BIB}

# Bibliography {-}

<div id="refs"></div>

\singlespace
\newpage
\fancypagestyle{APP}{%
    \fancyhead[R]{}
    \fancyhead[L]{}
}

\thispagestyle{APP}

# Appendix {-}
